---
title: "fars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Package for Summarizing and Plotting US Traffic Accidents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# A Package for Summarizing and Plotting US Traffic Accidents

This package is an assigment of Coursera course 'Building an R package', and has functions for summarizing and plotting maps from US traffic accidents data. Data is received from US National Highway Traffic Safety Administration's Fatality Analysis Reporting System, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. Instructions for data access here:  https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(fars)
library(dplyr)

```

```{r load-example-data, include=FALSE}
system.file("extdata", "accident_2013.csv.bz2",
            package = "fars")
system.file("extdata", "accident_2014.csv.bz2",
            package = "fars")
system.file("extdata", "accident_2015.csv.bz2",
            package = "fars")

```


## Functions

Package has fife functions for US accidents data summarizing and plotting:

* Main functions
  * `fars_summarize_years()`: summarizes accidents data 
  * `fars_map_state()`: draws map of accident locations by state and year  

* Helpers (used by main functions)
  * `fars_read()`: reads csv data
  * `make_filename()` : makes a filename by adding year to default data file 'accident_YYYY.csv.bz2'
  * `fars_read_years()`: returns accidents data by year and month  

## Summarizing yearly accidents
Select one or several years for summaries, function will return a tibble with monthly number of accidents each year.


```{r, iclude = FALSE, warning = FALSE, message=FALSE}
fars_summarize_years(c(2013, 2014, 2015))
```


## Plot maps with accidents
Accident locations can be plotted by state and year.
```{r, iclude = FALSE, warning = FALSE, message=FALSE}
fars_map_state(41, 2014)
```
